GVN-Hoist: Hoisting Computations from Branches

### 2016-08-17 hoist.tex
profitability

\subsection{Checking for profitability}
\label{subsec:profitability}
Establishing a set of cost models tune for performance. Like decreasing code
size, minimizing spills etc.

\subsubsection{Hoisting scalars}
Care must be taken in case of hoisting scalars too far, as that may increase
register pressure and result in spills. For example hoisting a scalar past a
call (Section-\ref{cost:across-calls}). In our current implementation we hoist
scalars past a call only when optimizing for code-side (-Os). Another way to
mitigate this problem is be to reinstantiate (rematerialize) the computation
after a call (may be as a different optimization pass).

@@

\subsection{Making the transformation}
Once all the legality checks and profitability checks are satisfied for a set of
identical instructions, they are suitable candidates for hoisting. A copy of the
computation is inserted at the hoisting point along with any instructions which
needed to be rematerialized. Thereafter, all the computations made redundant by
the new copy are removed. After that SSA form is restored by updating the
intermediate representation (IR) to reflect changes. After one iteration of
algorithm runs through the entire function, it creates more oppportunities for
\italics{higher ranked} computations \cite{rosen1988global}. Currently, we rerun
the algorithm until there are no more instructions left to be
hoisted. Obviously, this is not the most optimal approach and can be improved by
ranking the computations \cite{rosen1988global}.

>>>>

\subsection{Checking for profitability}
\label{subsec:profitability}
After the legality checks have passed, we check for profitability of hoisting.
That takes into account the impact code-hoisting would have on various
parameters that affects runtime performance e.g., impact on live-range, gain in
the code size.  We have established a set of cost models \ref{sec:cost-models}
for each parameter and tuned them for performance against representative
benchmarks.

\subsubsection{Hoisting scalars}
Since scalars are the majority of instructions which are hoisted, we pay special
attention in case of hoisting scalars too far, as that may increase register
pressure and result in spills. For example hoisting a scalar past a call
(Section-\ref{cost:across-calls}).  In our current implementation we hoist
scalars past a call only when optimizing for code-side (-Os). Ideally, a later
stage of live-range splitting pass should split the live-ranges for optimal
performance, however, that is not the case with llvm as we have found
regressions when scalars are hoisted too far \ref{sec:cost-models}. Another way
to mitigate this problem is be to reinstantiate (rematerialize) the computation
after a call (may be as a different optimization pass).

@@

\subsection{Making the transformation}
Once all the legality checks and profitability checks are satisfied for a set of
identical instructions, they are suitable candidates for hoisting. A copy of the
computation is inserted at the hoisting point along with any instructions which
needed to be rematerialized. Thereafter, all the computations made redundant by
the new copy are removed. After that SSA form is restored by updating the
intermediate representation (IR) to reflect changes. After one iteration of
algorithm runs through the entire function, it creates more oppportunities for
\emph{higher ranked} computations \cite{rosen1988global}. Currently, we rerun
the algorithm until there are no more instructions left to be
hoisted. Obviously, this is not the most optimal approach and can be improved by
ranking the computations \cite{rosen1988global}.
